---
title: "Reburns_Regen_analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(here)
theme_set(theme_cowplot())
options(scipen = 9999)
se <- function(x) sqrt(var(x)/length(x))

library(sjPlot)
library(MASS)
library(logNormReg)
```

# Effects of short-interval fires continue to accumulate, overwhelming the effects of local resiliency

This file includes code used to clean up, organize and analyze data on regeneration density and basal area of reburned plots in Interior Alaska. Raw data files are saved as "_raw.csv", and as much work as possible is contained within the code itself. 

## Regeneration density

Density here is density of trees and seedlings on a scale of stems per hectare for each plot. 

Code cleans up DBH, adds division (conifer / deciduous) to species, adds expansion factors specific to each plot (both dbh and seedlings were counted across different sub-samples for each plot, so scaling to ha is a bit more complicated).

```{r clean up of dbh data}
# dbh <- read.csv(here("data/dbh_raw.csv"))
# 
#   # merging salix # so many willows, so little time
#   dbh$SPP[dbh$SPP == "SAGL"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_3"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_4"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_5"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_6"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_7"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_8"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SA_?"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SADE"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SAPU"] <- "SALIX"
#   dbh$SPP[dbh$SPP == "SAGL_R"] <- "SALIX"
#   
#    # dropping unknown
#      dbh <- dbh %>%
#        filter(SPP != "UNKNOWN") # only a few unknown species present at Dalton site
#   
#  # adding conifer and deciduous divisions
#     dbh$DIV[dbh$SPP == "PIME"] <- "c"
#     dbh$DIV[dbh$SPP == "BENE"] <- "d"
#     dbh$DIV[dbh$SPP == "POTR"] <- "d"
#     dbh$DIV[dbh$SPP == "POBA"] <- "d"
#     dbh$DIV[dbh$SPP == "SALIX"] <- "d"
#     dbh$DIV[dbh$SPP == "ARCTO"] <- "d"
#     dbh$DIV[dbh$SPP == "ALCR"] <- "d"
# 
# # adding expansion factors
#     # assuming a plot is 1/25 of a ha, so half a plot is 1/50 of a ha.
#   dbh$EXP_FACT[dbh$QUAD == 2] <- 50
#   dbh$EXP_FACT[dbh$QUAD == 1] <- 100
#   dbh$EXP_FACT[dbh$QUAD == .1] <- 1000
#   dbh$EXP_FACT[dbh$QUAD == .2] <- 500
# 
#  write.csv(dbh, here("data/dbh.csv"), row.names = F)
```

The next chunk brings in the altered dbh csv file, drops dead trees, sums the count of each species by plot and adds the division and expansion factors back in. The result is a file of densities of each plot according to species. 

```{r dbh sum, expansion factors and divisions}
dbh <- read.csv(here("data/dbh.csv"))

dbh$TREAT <- as.factor(dbh$TREAT)

  # dropping dead trees
    dbh <- dbh %>%
      filter(CANOPY > 0)
  # summing according to species
  dens <- dbh %>%
    group_by(SITE, TREAT, PLOT, SPP) %>%
    summarise(COUNT_PLOT = n()) %>%
    ungroup() %>%
    group_by(SITE, TREAT,PLOT) %>% 
    complete( SPP, fill = list(COUNT_PLOT = 0)) # adds zero-counts back in for species
  
  # adding divisions back in
    dens$DIV <- NA
    dens$DIV[dens$SPP == "PIME"] <- "c"
    dens$DIV[dens$SPP == "BENE"] <- "d"
    dens$DIV[dens$SPP == "POTR"] <- "d"
    dens$DIV[dens$SPP == "POBA"] <- "d"
    dens$DIV[dens$SPP == "SALIX"] <- "d"
    dens$DIV[dens$SPP == "ARCTO"] <- "d"
    dens$DIV[dens$SPP == "ALCR"] <- "d"
    dens$DIV <- as.factor(dens$DIV)
    
  # adding expansion factor back
    test <- dbh %>%
    group_by(SITE, TREAT, PLOT) %>%
    summarise(EXP_FACT = max(EXP_FACT)) %>%
      slice(rep(1:n(), each = 7)) %>% # 7 species at each plot, so each plot is 7 rows
      ungroup()
  
    dens$EXP_FACT <- test$EXP_FACT # adding expansion factors back in

# scaling up
    dens$COUNT_ha <- dens$COUNT_PLOT * dens$EXP_FACT 
```

The following code brings in the seedling data file, adds expansion factors specific to seedlings (since they were counted separately from dbh), scales seedling counts to stems/ha and combines the files of species dbh density and species seedling density. No white spruce existed above dbh height at any of the plots, but some seedlings occurred, so I also add a row of zero counts for white spruce to each dbh plot. The result from the entire code chunk is a datafile ("density.csv") containing counts of seedlings and trees of each species for each plot, along with their respective expansion factors and counts per hectare. 

```{r seed sum, expansion fact, divisons and adding with dbh}
seed <- read.csv(here("data/seedling_count_raw.csv"))

# adding expansion factors
  seed$EXP_FACT <- NA
  seed$EXP_FACT[seed$M2 == 100] <- 100
  seed$EXP_FACT[seed$M2 == 10] <- 1000
  seed$EXP_FACT[seed$M2 == 20] <- 500
  seed$EXP_FACT[seed$M2 == 9] <- (10000/9)
  
# scaling up
  seed$COUNT_ha <- seed$EXP_FACT * seed$TOTAL_COUNT
  
# combining seedling and dbh
  seed <- seed %>% 
    group_by(SITE, TREAT, PLOT) %>%
    arrange(match(SPP, c("ALCR", "ARCTO", "BENE", "POBA", "POTR", "SALIX", "PIGL", "PIME")),
            .by_group = T) # arranges species in the same order within plots
                           # also orders plots in same order within sites
                           # need to do to combine with dbh in a sec

  # creating PIGL rows 
  # no white spruce at dbh height on plots, but need to add to combine with seedling file
  test <- dens %>%
    dplyr::select(SITE, TREAT, PLOT, EXP_FACT) %>% # need the dplyr because MASS package conflicts with select()
    distinct() # creates file where each row is a plot, maintains expansion factors/site/treat
  test$SPP <- as.factor("PIGL") # adds species
  test$COUNT_ha <- 0 # adds zero count
  test$COUNT_PLOT <- 0
  test$DIV <- as.factor("c")
  dens <- rbind(dens, test) # puts back into dbh file 
  
   dens <- dens %>%
    group_by(SITE, TREAT, PLOT) %>% # same function as above # specifies order of species
    arrange(match(SPP, c("ALCR", "ARCTO", "BENE", "POBA", "POTR", "SALIX", "PIGL", "PIME")),
            .by_group = T) %>%
     rename(TREE_COUNT_PLOT = COUNT_PLOT, TREE_COUNT_HA = COUNT_ha,
            TREE_EXP_FACT = EXP_FACT) # renames tree counts so we can combine with seedling
   
  # combining files
  dens$SEED_COUNT_PLOT <- seed$TOTAL_COUNT
  dens$SEED_COUNT_HA <- seed$COUNT_ha
  dens$SEED_EXP_FACT <- seed$EXP_FACT

    # making sure number of fires is a factor
    dens$TREAT <- as.factor(dens$TREAT)
    
    # creating total count of both trees and seedlings
    dens$COUNT_HA <- dens$TREE_COUNT_HA + dens$SEED_COUNT_HA
    
 #  write.csv(dens, here("data/density.csv"), row.names = F)
```

Finally, this last chunk brings in the values of slope, solar radiation and elevation of each site, to be used as fixed effects in the later model. 

```{r bringing in site attributes as covariates}
    # removing unburned sites
    dens <- dens[dens$TREAT !=0,]
    dens$TREAT <- as.factor(dens$TREAT)

    # looking at format of data, first three rows
    dens[1:3,]
    
# pulling in site attribute variables
    attrib <- read.csv(here("data/site_attrib.csv"))

    # no unburned sites used here
    attrib <-attrib[attrib$TREAT != 0,]
 
    # getting order of sites in dens
    order <- as.vector(unique(dens$PLOT)) # don't worry that its 42, remember this doesn't include unburned sites
    attrib <- attrib %>%
      group_by(SITE, TREAT) %>%
      arrange(match(PLOT,order)) %>% # matches the same order of plots as density file
      slice(rep(1:n(), each = 8)) # replicates each value by 8, since there's 8 rows per plot (8 species)
    # adding in attributes to density file
    dens$SLOPE <- attrib$SLOPE
    dens$ELEV <- attrib$ELEVATION
    dens$SOLAR <- attrib$SOLAR
```

The piece of code that follows is the analysis of conifer density. The first half of the script uses base R to subset for just conifers, and then sums density of white and black spruce (the 2 conifers present at any site) before bringing the other variables back in. I checked the performance of negative binomial, poisson and quasipoisson distributions before deciding to go with the negative binomial, based on the fit of the residuals in the qq plot.I then added in slope and solar radiation as fixed effects, and re-checked the diagnostics. 

```{r conifer density analysis}
dens$TREAT <- as.factor(dens$TREAT)

# Different way of subsetting, base R  
  CONIF_dens <- subset(dens, dens$DIV == "c")

#the plots are duplicated with one entry per conifer species, so this reduces it to one value per plot
  cnt <- aggregate(CONIF_dens$COUNT_HA, list(CONIF_dens$PLOT), FUN = sum)  	#adds up the count per ha for each row within a plot
  trt <- aggregate(CONIF_dens$TREAT, list(CONIF_dens$PLOT), FUN = mean)	#gets the treatment number (mean works because they are the same
  CONIF_dens_new <- cbind(cnt, trt[,2])						#combines
  
  temp <- merge(x=CONIF_dens_new, y=CONIF_dens, by.x="Group.1", by.y="PLOT", all.x=F, all.y=F)	#adding in the site in a clumsy way but it was quick
  CONIF_dens_new <- temp[duplicated(temp[,1]),]					#merging and removing
  CONIF_dens <- CONIF_dens_new %>%	
    rename(PLOT = Group.1, COUNT = x) %>% # renaming columns
    dplyr::select(SITE, TREAT, PLOT, COUNT, SLOPE, ELEV, SOLAR) # selecting only the things we'll use
  hist(CONIF_dens$COUNT)									#check
  
  mC_D.nb <- 	glm.nb(COUNT ~ TREAT + SITE + (TREAT*SITE), link = log, data = CONIF_dens) 
  mC_D    <- 	glm(COUNT ~ TREAT + SITE + (TREAT*SITE), family = poisson(link = "log"), data = CONIF_dens)
  mC_D_qp <- 	glm(COUNT ~ TREAT + SITE + (TREAT*SITE), family = quasipoisson(link = "log"), data = CONIF_dens) 
  
  # residuals
  par(mfrow=c(3,4))
  plot(mC_D, main="Poisson")
  plot(mC_D_qp, main="Quasi-poisson")
  plot(mC_D.nb, main="NB")
  
  # parameters
  summary(mC_D)
  summary(mC_D_qp)
  summary(mC_D.nb) # AIC 931
  
  #note the prediction vs. observed isn't great, but that's not the point here
  dev.off()
  plot(predict(mC_D.nb, type="response",se.fit=T)$fit,CONIF_dens$COUNT)	#plots the exponentiated prediction back
  abline(0,1, col="red")								#one to one line
  
  #the model doesn't do bad at predicting areas with zero trees (treatment effect) are zero, it's the variability
  #in the regen where there is regen (treatments 1 and 2 primarily) that must be explained by other factors.
  #That's clear from this prediction plot and where the major errors are, but also in the standard errors.
  
  plot(predict(mC_D.nb, type="response",se.fit=T)$fit, predict(mC_D.nb, type="response",se.fit=T)$se.fit, ylab="SE", xlab="Prediction")
  abline(0,1, col="red")
  
  #bringing in other covariates

  #to do so, adding them into the glm above, but then check the diagnostics again of course
    mC_D.nb <- 	glm.nb(COUNT ~ TREAT + SITE + (TREAT*SITE), link = log, data = CONIF_dens)
    mC_D.nb.slope <- 	glm.nb(COUNT ~ TREAT + SITE + SLOPE  + (TREAT*SITE),
                             link = log, data = CONIF_dens)

    par(mfrow=c(2,4))
    plot(mC_D.nb, main="NB")
    plot(mC_D.nb.slope, main="Slope")

  summary(mC_D.nb) # AIC 924
  summary(mC_D.nb.slope) # AIC 926
  
  confint(mC_D.nb)

```

The next piece of code conducts the same process, but for deciduous density. 

```{r deciduous density analysis}
# Different way of subsetting, base R  
  DECID_dens <- subset(dens, dens$DIV == "d")

  # the plots are duplicated with one entry per species, so this reduces it to one value per plot
  cnt <- aggregate(DECID_dens$COUNT_HA, list(DECID_dens$PLOT), FUN = sum)  	#adds up the count per ha for each row within a plot
  trt <- aggregate(DECID_dens$TREAT, list(DECID_dens$PLOT), FUN = mean)		#gets the treatment number (mean works because they are the same
  DECID_dens_new <- cbind(cnt, trt[,2])						#combines
  
  temp <- merge(x=DECID_dens_new, y=DECID_dens, by.x="Group.1", by.y="PLOT", all.x=F, all.y=F)	#adding in the site in a clumsy way but it was quick
  temp2 <- temp[duplicated(temp[,1]),]										#merging and removing
  temp3 <- temp2[duplicated(temp2[,1]),]        # have to loop through, since there's so many more species
  temp4 <- temp3[duplicated(temp3[,1]),]        # absolutely sure there's a faster way to do this, just wanted to get it done
  temp5 <- temp4[duplicated(temp4[,1]),]
  DECID_dens_new <- temp5[duplicated(temp5[,1]),]				
  rm(temp,temp2, temp3, temp4, temp5, cnt, trt) 				# cleaning up workspace
  DECID_dens <- DECID_dens_new %>%	
    rename(PLOT = Group.1, COUNT = x) %>% # renaming columns
    dplyr::select(SITE, TREAT, PLOT, COUNT, SLOPE, ELEV, SOLAR) # selecting only the things we'll use							#back to original name so code below works
  dev.off()
  hist(DECID_dens$COUNT)									#check
  
  mD_D.nb <- 	glm.nb(COUNT ~ TREAT + SITE + (TREAT*SITE), link = log, data = DECID_dens) 
  mD_D <- 		glm(COUNT ~ TREAT + SITE + (TREAT*SITE), family = poisson(link = "log"), data = DECID_dens) 
  mD_D.qp <- 	glm(COUNT ~ TREAT + SITE + (TREAT*SITE), family = quasipoisson(link = "log"), data = DECID_dens) 
  
  # residuals
  par(mfrow=c(3,4))
  plot(mD_D, main="Poisson")
  plot(mD_D.qp, main="Quasi-poisson")
  plot(mD_D.nb, main="NB")
  
  sqrt(mean(mD_D.nb$residuals^2))
  sqrt(mean(mD_D$residuals^2))
  sqrt(mean(mD_D.qp$residuals^2))

  # parameters
  summary(mD_D)
  summary(mD_D_qp)
  summary(mD_D.nb) # AIC 1087
  
  dev.off()
  plot(predict(mD_D.nb, type="response",se.fit=T)$fit,DECID_dens$COUNT)	#plots the exponentiated prediction back
  abline(0,1, col="red")								#one to one line
  
  plot(predict(mD_D.nb, type="response",se.fit=T)$fit, 
       predict(mD_D.nb, type="response",se.fit=T)$se.fit, ylab="SE", xlab="Prediction")

  #bringing in other covariates
  #adding them into the glm above
  mD_D.nb <- 	glm.nb(COUNT ~ TREAT + SITE + (TREAT*SITE), link = log, data = DECID_dens)
  mD_D.nb.slope <- 	glm.nb(COUNT ~ TREAT + SITE + SLOPE + (TREAT*SITE),link = log, data = DECID_dens)
  
  dev.off()
  par(mfrow=c(5,4))
  plot(mD_D.nb, main="NB")
  plot(mD_D.nb.slope, main="Slope")
  
  summary(mD_D.nb) # AIC 1087
  summary(mD_D.nb.slope) # AIC 1089
  
  confint(mD_D.nb)
```


Finally, the code below pulls out specific values of density used in the text of the manuscript. 

```{r density values for text}
# getting values 
dens <- read.csv(here("data/density.csv"))

treat_dens <- dens %>%
  group_by(SITE, TREAT, DIV) %>%
  summarise(AV = mean(COUNT_HA), SE = se(COUNT_HA))

treat_dens$SE <- round(treat_dens$SE, digits = 0)
treat_dens$AV <- round(treat_dens$AV, digits = 0)

treat_dens %>%
  filter(SITE == "DALTON") %>%
  filter(DIV == "c")

stee_con_dens <- treat_dens %>%
  filter(SITE == "STEESE") %>%
  filter(DIV == "c")

treat_dens %>%
  filter(SITE == "DALTON") %>%
  filter(DIV == "d")

treat_dens %>%
  filter(SITE == "STEESE") %>%
  filter(DIV == "d")
```

## Basal area


```{r basal area bring in covariates}
# bringing in data
  ba <- read.csv(("ba.csv"), stringsAsFactors = F)
 
  # removing unburned sites
  ba <- ba[ba$TREAT != 0,]
  
  # looking at format of data, first three rows
  ba[1:3,]
  
   DECID_BA <- subset(ba, ba$DIV == "d")

  DECID_BA$QUAD[DECID_BA$QUAD == 2] <- 50	#assuming a plot is 1/25 of a ha, so half a plot is 1/50 of a ha.
  DECID_BA$QUAD[DECID_BA$QUAD == 1] <- 100
  DECID_BA$QUAD[DECID_BA$QUAD == .1] <- 1000
  DECID_BA$QUAD[DECID_BA$QUAD == .2] <- 500
 
  #calc basal area
  DECID_BA$BA <- (pi * (DECID_BA$DBH/2)^2)/10000	#convert to ba and cm2 to m2. Just pi*r2
  DECID_BA$BA_ha <- DECID_BA$BA * DECID_BA$QUAD		#multiply by expansion factor 
  
DECID_BA <- DECID_BA %>%
    group_by(SITE, TREAT, PLOT, DIV) %>%
    summarise(BA = sum(BA_ha))

  #histogram
  hist(DECID_BA$BA)

  # pulling in site attribute variables
    attrib <- read.csv(here("data/site_attrib.csv"))
    
    # no unburned
    attrib <-attrib[attrib$TREAT != 0,]
    
    # getting order of sites in dens
    order <- as.vector(DECID_BA$PLOT) # don't worry that its 42, remember this doesnt include unburned
    attrib <- attrib %>%
      group_by(SITE, TREAT) %>%
      arrange(match(PLOT,order)) 
    # adding in attributes to density file
    DECID_BA$SLOPE <- attrib$SLOPE
    DECID_BA$ELEV <- attrib$ELEVATION
    DECID_BA$SOLAR <- attrib$SOLAR
```


```{r basal area analysis}
DECID_BA$TREAT <- as.factor(DECID_BA$TREAT)

  mD_BA.log <- lm(log(BA)  ~ TREAT + SITE + (TREAT*SITE), data = DECID_BA)						#models E[log(y)]
  mD_BA.ga <- glm(BA  ~ TREAT + SITE + (TREAT*SITE), family = gaussian(link = "log"), data = DECID_BA)	#models log[E(y)]
  mD_BA.gamma <- glm(BA  ~ TREAT + SITE + (TREAT*SITE), family = Gamma(link = "log"), data = DECID_BA)
  
    #another option, look up lognormal distributions
  BA_logn <- lognlm((BA)  ~ TREAT + SITE + (TREAT*SITE), data = DECID_BA)

  par(mfrow=c(4,4))
  plot(mD_BA.log, main= "Logged")
  plot(mD_BA.ga, main="Guassian")
  plot(mD_BA.gamma, main="Gamma")


  #choosing a distribution is often best done with an understanding of the relationship between variance and mean
  #Gaussian has constant variance; gamma and lognormal do not.
  #heteroskedastic
  par(mfrow=c(1,4))
  plot((DECID_BA$BA),residuals(mD_BA.log),main="log"); abline(h=0)
  plot((DECID_BA$BA),residuals(mD_BA.ga), main="Gaussian"); abline(h=0)	#pretty clear Gaussian is not great.
  plot((DECID_BA$BA),residuals(mD_BA.gamma), main="Gamma"); abline(h=0)
  plot((DECID_BA$BA),residuals(BA_logn), main="Lognormal"); abline(h=0)

  # parameters
  summary(mD_BA.log)
  summary(mD_BA.ga) # AIC 133
  summary(mD_BA.gamma) # AIC 58
  summary(BA_logn)

  #Compare root mean squared error
  sqrt(mean(mD_BA.log$residuals^2)) # 0.80
  sqrt(mean(mD_BA.ga$residuals^2)) # 0.82
  sqrt(mean(mD_BA.gamma$residuals^2)) # 0.82
  sqrt(mean(BA_logn$residuals^2)) # 1.03

  par(mfrow=c(1,4)	)
  plot(exp(predict(mD_BA.log)),DECID_BA$BA, main="logged LM")	#plots the exponentiated prediction back
  abline(0,1)
  plot((predict(mD_BA.ga, type="response")),DECID_BA$BA, main="Guassian")	#plots the exponentiated prediction back
  abline(0,1)
  plot(predict(mD_BA.gamma, type="response"),DECID_BA$BA, main="Gamma with log link")	#plots the exponentiated prediction back
  abline(0,1)
  plot(BA_logn$fitted.values,DECID_BA$BA, main = "lognormal")	#plots the exponentiated prediction back
  abline(0,1)

  # adding in covariates
  mD_BA.gamma <- glm(BA  ~ TREAT + SITE + (TREAT*SITE), family = Gamma(link = "log"), data = DECID_BA)
  mD_BA.gamma.slope <- glm(BA  ~ TREAT + SITE + SLOPE + (TREAT*SITE), family = Gamma(link = "log"), data = DECID_BA)
 
    par(mfrow=c(4,4))
  plot(mD_BA.gamma, main= "Gamma")
  plot(mD_BA.gamma.slope, main="Slope")
  
  summary(mD_BA.gamma) # AIC 58.2
  summary(mD_BA.gamma.slope) # AIC 58.8
  
  confint(mD_BA.gamma)
```


```{r basal area values for text}
treat_ba <- ba %>%
  group_by(SITE, TREAT, DIV) %>%
  summarise(AV = mean(BA_ha), SE = se(BA_ha))

treat_ba %>%
  filter(SITE == "DALTON") %>%
  filter(DIV == "d")

treat_ba %>%
  filter(SITE == "STEESE") %>%
  filter(DIV == "d")
```

