---
title: "Soil Analysis"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Adventitious roots, organic layer and exposed mineral soil

Working with 3 forms of soil data here: organic layer depth, percent cover of exposed mineral soil and depth to adventitious roots. 

```{r setting up}
  library(here)
  library(ggplot2)
  library(cowplot) ; theme_set(theme_cowplot())
  library(dplyr)
```

# Adventitous roots

Plotting depth to adventitious roots, product is a supplemental figure. 

```{r adventitous roots plot}
adven_root <- read.csv(here("data/adventaguous_roots.csv"))

adven_root_plot <- ggplot(adven_root, 
                          aes(x = factor(TREAT),
                              y = DEPTH, 
                              fill = SITE)) + 
  geom_boxplot() + 
  labs(x = "Number of Fires", y = "Depth (cm)",
       title = "Soil Consumption") + 
  scale_fill_manual(values = c("#d9d9d9", "#ffffff"), 
                    name = "Site Type", 
                    labels = c("Upland", "Lowland")) + 
  theme(plot.title = element_text(hjust = 0),
        legend.position = "none") + xlim("1", "2")

adven_root_plot
# save_plot("adven_root.png", adven_root_plot, base_aspect_ratio = 1.5)
```

# Organic layer depth

```{r organic layer plot}
org_depth <- read.csv(here("data/org_depth.csv"))

 org_depth_plot <- ggplot(org_depth, aes(x = factor(TREAT),
                                         y = ORG_DEPTH, 
                                         fill = SITE)) + 
    geom_boxplot()   + 
    scale_fill_manual(name = "Site Type",
                      values = c("#d9d9d9", "#ffffff"), 
                      labels = c("Upland", "Lowland")) + 
    labs(title = "Organic Layer Depth", 
         x = "Number of Fires", y = "Organic layer depth (cm)") + 
    theme(plot.title = element_text(hjust = 0), 
          legend.position =  "none") 
 
org_depth_plot
  
# save_plot("org_depth.png", org_depth_plot, base_aspect_ratio = 1.5)
```

# Exposed mineral soil

```{r exposed mineral soil}
exposed_min <- read.csv(here("data/exposed_min.csv"))

expose_min_plot <- ggplot(exposed_min, 
                          aes(x = factor(TREATMENT),
                              y = MINERAL,
                              fill = SITE)) + 
    geom_boxplot() + 
    scale_fill_manual(name = "Site Type",
                      values = c("#d9d9d9", "#ffffff"), 
                      labels = c("Upland", "Lowland")) + 
    labs(title = "Exposed Mineral Soil", 
         x = "Number of Fires", y = "% of Mineral Soil Exposed") + 
    theme(plot.title = element_text(hjust = 0))

expose_min_plot
```

# Figure 7

Produces figure 7, which includes both exposed mineral soil and organic layer depth. 

```{r combined}
substrate <- plot_grid(org_depth_plot, expose_min_plot,
          nrow = 1, ncol = 2,
          rel_widths = c(0.9,1.3), rel_heights = c(1, 1.5),
          labels = c("A.", "B."))

save_plot("substrate.png", substrate, nrow = 1, ncol = 2,
          base_aspect_ratio = 1.5)
```

# getting numbers
## adven root
```{r getting mins and maxs uplands}
upland <- adven_root[adven_root$SITE == "DALTON",]

summary(upland$DEPTH[upland$TREAT == 1])

summary(upland$DEPTH[upland$TREAT == 2])
24/10
```

```{r getting mins and maxs lowland}
lowland <- adven_root[adven_root$SITE == "STEESE",]

summary(lowland$DEPTH[lowland$TREAT == 1])
low1 <- lowland$DEPTH[lowland$TREAT == 1]
low2 <- lowland$DEPTH[lowland$TREAT == 2]
t.test(low1,low2)
```

## organic layer

```{r getting maximums and minimums}
org_treat <- org_depth %>%
  group_by(SITE, TREAT) %>%
  summarise(AV_DEPTH = mean(ORG_DEPTH))

org_treat$AV_DEPTH <- round(org_treat$AV_DEPTH, digits = 2)
```

